\documentclass[letterpaper,10pt,serif,draftclsnofoot,onecolumn,compsoc,titlepage]{article}%this needs to change to IEEEtran but there is a bug

\usepackage{graphicx}                                        
\usepackage{amssymb}                                         
\usepackage{amsmath}                                         
\usepackage{amsthm}                                          

\usepackage{alltt}                                           
\usepackage{float}
\usepackage{color}
\usepackage{url}

\usepackage{balance}
\usepackage[TABBOTCAP, tight]{subfigure}
\usepackage{enumitem}
\usepackage{pstricks, pst-node}

\usepackage{geometry}
\geometry{margin=.75in}
\usepackage{hyperref}
\usepackage{tikz}
%\usetikzlibrary{shapes, positioning, calc}
\colorlet{lightgray}{gray!20}
\usepackage{caption}
\usepackage{listings}
%\usepackage[utf8]{inputenc}
%pull in the necessary preamble matter for pygments output
\input{pygments.tex}

%% The following metadata will show up in the PDF properties
% \hypersetup{
%   colorlinks = false,
%   urlcolor = black,
%   pdfauthor = {\name},
%   pdfkeywords = {cs311 ``operating systems'' files filesystem I/O},
%   pdftitle = {CS 311 Project 1: UNIX File I/O},
%   pdfsubject = {CS 311 Project 1},
%   pdfpagemode = UseNone
% }

\parindent = 0.0 in
\parskip = 0.1 in
\title{Homework 1 Write Up}
\author{Shannon Ernst, Keith Stirby, Tanner Cecchetti \\ 21 April 2017 \\ CS 444 Spring 2017 \\ Group 10-01}
\begin{document}
\maketitle
\begin{abstract}
This is the abstract
\end{abstract}
\newpage
\section{Log of Commands}
\section{qemu Flag Explanation}
\section{Concurrency Write Up}
\subsection{Main Point of the Assignment}
The main point of the assignment was to introduce the idea of 
concurrency and refresh on threads. The basics of the assignment was
to spawn a provided number of threads which would arbitrarily be assigned
 producer or consumer status. These threads would then operate on the same
 buffer space adding and removing jobs. This simple simulation allowed 
 the student to become comfortable with pthreads. 
\subsection{Design Decisions}
There were very few concrete requirements for this assignment. The few we 
did have were:
\begin{itemize}
\item Producers create an object containing a random number and a random 
wait time ranging from 2-9 seconds.
\item Producers add their object to a buffer so long as there are fewer 
than 32 items in the buffer after a random wait time of 3-7 seconds.
\item Consumers remove jobs from the buffer so long as there are more than 
zero in the buffer
\item Consumers read the wait time on a given item and wait for that amount 
of time before printing the item's random value.
\item Randomization technique is determined by the system and will either 
use rdrand or the Mersenne Twister.
\end{itemize}
Beyond these requirements, we were left to design the rest. We decided to 
begin with the synchronization technique. This assignment needs to have 
all of the threads running at the same time on the same buffer. We decided 
to use mutexes to control the locking and unlocking of the buffer. A 
producer will take a lock on the buffer to check if it is full. If it is 
the lock is released immediately. If the buffer is not full, the producer 
maintains the lock and will create an object. When an object is created 
the program checks if it can use rdrand and then makes a selection on 
which randomization technique to use. The item is then added to the current 
counter position in the buffer and then the counter is incremented. The 
buffer is an array which acts as a stack: the first item will be the last 
item removed. We made this decision for simplicity's sake and recognize 
that a better scheduler would be a first in first out structure which 
would have just required a rotation of adding to the array, keeping two 
index counters or pointers to keep track of where to consume from and 
where to produce to. Similar to producers, consumers take a lock on the 
buffer and check if it is empty. If it is empty, the consumer releases the 
lock. If there is an item in the buffer the consumer removes it, 
decrementing the counter, and reads the wait time. The consumer calls the 
sleep command for that amount of time. When the consumer wakes up it prints 
the value. The program continues until a kill command is received from the 
keyboard. The number of threads is received from the command line. 
\subsection{Testing}
We did incremental testing as we developed. The trickiest thing to test was 
the randomization techniques. We began by testing the rdrand locally on a 
new MAC laptop which successfully invoked the randomization. Having 
confirmed that our program would run the rdrand function, we moved onto 
the os-class server and implemented the Mersenne Twister. We tested with a 
variety of threads on the os-class maxing out at 4, printing out traces of 
the functions being invoked. 
\subsection{Take Aways}
Coming into this assignment, there were some reservations about implementing
 pthreads. This assignment has shown us that pthreads is not as difficult as
 we feared. We also learned how to correctly lock and unclock implicitly 
 shared data. 
\section{Version Control Log}
\section{Work Log}
Tuesday, 11 April, 2017, 10:00 am: we met our group for the first time in
 recitation, we attempted to set up the vm and kernal boot in recitation
  unsuccessfully \\
Wednesday, 12 April 2017, 6:00 pm: we met to successfully boot the vm and 
the kernal. We were able to get the vm to work but could not get the kernal 
to boot. We were going to ask questions later about it.\\
Wednesday, 19 April, 2017, 6:00 pm: we met and coded the concurrency 
assignment.\\ 
Thursday, 20 April, 2017, 9:00 pm: Shannon created the latex file, the 
make file and wrote the concurrency write up section. 


\end{document}
